Investigation of ant cuticle dataset using image texture analysis methods

### 2022-07-29 paper/main.tex
add subsections, update images

This paper is organized as below:  Section 2 reviews the sculpturing
identification in ants, and image texture analysis and insect classification.
Section 3 describes the dataset preparation and image texture analysis methods
used in our experiments. Section 4 gives the experimental results, visualization
of both raw and processed images and analysis. The conclusion then follows.

\section{Related work on sculpturing identification in entomology, image texture
  analysis, and insect classification}

% \subsection{Sculpturing identification}

@@

% \subsection{Texture analysis}

@@

% \subsection{Insect classification}

@@

    \begin{subfigure}{\segmentedsubwidth}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/sr103.png}
    \end{subfigure}
    \begin{subfigure}{\segmentedsubwidth}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/sr111.png}
    \end{subfigure}
    \begin{subfigure}{\segmentedsubwidth}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/ss105.png}
    \end{subfigure}
    \begin{subfigure}{\segmentedsubwidth}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/ss107.png}
    \end{subfigure}

    \begin{subfigure}{\segmentedsubwidth}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/heat103.png}
    \end{subfigure}
    \begin{subfigure}{\segmentedsubwidth}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/heat111.png}
    \end{subfigure}
    \begin{subfigure}{\segmentedsubwidth}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/heat105.png}
    \end{subfigure}
    \begin{subfigure}{\segmentedsubwidth}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/heat107.png}
    \end{subfigure}
    \caption{Raw images are shown with corresponding segmented outputs from
        K-views algorithm. The second row shows the results of using size of
        15x15 patches. The third row shows the results of using size of 19x19
        patches.}
    \label{fig:kviews_segmented}
\end{figure}

\FloatBarrier Some classified results from the K-views are shown in Figure
\ref{fig:kviews_segmented}. A training set of 500 patches for background, 500
patches for rough, and 500 patches for smooth are applied with patch sizes of
15x15 and 19x19. These patches are extracted from two rough and two smooth
images. The K-views algorithm converges early although we set the maximum number
of iterations to be 1000.

\FloatBarrier
\section{Conclusion}
Ant cuticle texture presumably has some function, but without the proper tools,
evaluating the function based on thousands of species is infeasible. We have
shown in this work that a deep learning approach and classical image texture
analysis methods can be used to automatically categorize ants based on their
cuticle texture, therefore supporting research on the evaluation of the function
in future work. Our categorization system is novel in the field of automated
insect identification due to the broad number of species captured by it.
Additionally, a model that is pre-trained on a diverse image task such as ResNet
can be transferred to our domain of texture analysis. However, a deep learning
algorithm created specifically for the domain of texture analysis had the best
results in our experiments. In future work, we will continue to explore texture
classes present in the dataset which are not captured by the binary class
system. All code is publicly available on GitHub
(https://github.com/ngngardner/cuticulus).

>>>>

This paper is organized as below:  Section 2 reviews the microsculpturing
identification in ants, and image texture analysis and insect classification.
Section 3 describes the dataset preparation and image texture analysis methods
used in our experiments. Section 4 gives the experimental results, visualization
of both raw and processed images and analysis. The conclusion then follows.

\section{Related work}

In this section we review related work on microsculpturing identification, image
texture analysis, and insect classification.

\subsection{Microsculpturing identification}

@@

\subsection{Image texture analysis}

@@

\subsection{Insect classification}

@@

    \begin{subfigure}{\segmentedsubwidth}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/15/r103.png}
    \end{subfigure}
    \begin{subfigure}{\segmentedsubwidth}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/15/r111.png}
    \end{subfigure}
    \begin{subfigure}{\segmentedsubwidth}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/15/s105.png}
    \end{subfigure}
    \begin{subfigure}{\segmentedsubwidth}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/15/s107.png}
    \end{subfigure}

    \begin{subfigure}{\segmentedsubwidth}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/19/r103.png}
    \end{subfigure}
    \begin{subfigure}{\segmentedsubwidth}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/19/r111.png}
    \end{subfigure}
    \begin{subfigure}{\segmentedsubwidth}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/19/s105.png}
    \end{subfigure}
    \begin{subfigure}{\segmentedsubwidth}
        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/19/s107.png}
    \end{subfigure}
    \caption{Raw images are shown with corresponding segmented outputs from
        K-views algorithm. The second row shows the results of using size of
        15x15 patches. The third row shows the results of using size of 19x19
        patches. In the images, the color red represents the background class,
        green represents the smooth class, and the blue represent the rough
        class.}
    \label{fig:kviews_segmented}
\end{figure}

\FloatBarrier Some classified results from the K-views are shown in Figure
\ref{fig:kviews_segmented}. A training set of 550 patches for background, 455
patches for rough, and 493 patches for smooth are used with patch sizes of 15x15
and 19x19. These patches are extracted from two rough and two smooth images with
sizes of 608x608. The K-views algorithm converges early although we set the
maximum number of iterations to be 1000. These experimental results are
preliminary due to the minimal number of source images used and future work will
explore a more expansive approach to the problem.

\FloatBarrier
\section{Conclusion}
Ant cuticle texture presumably has some function, but without the proper tools,
evaluating the function based on thousands of species is infeasible. We have
shown in this work that a deep learning approach and classical image texture
analysis methods can be used to automatically categorize ants based on their
cuticle texture, therefore supporting research on the evaluation of the function
in future work. Our categorization system is novel in the field of automated
insect identification due to the broad number of species captured by it.
Additionally, a model that is pre-trained on a diverse image task such as ResNet
can be transferred to our domain of texture analysis \footnote{All code is
    publicly available on GitHub (https://github.com/ngngardner/cuticulus)}.
However, a deep learning algorithm created specifically for the domain of
texture analysis had the best results in our experiments. In future work, we
will continue to explore texture classes present in the dataset which are not
captured by the binary class system. We will also investigate application of
texture-based deep maching learning in image texture analysis.
